```{r, incluede=FALSE}
knitr::opts_chunk$set(cache = FALSE, warning = FALSE, message = FALSE, echo = FALSE, fig.topcaption = TRUE, fig.align = 'center')
options(knitr.kable.NA = ' ')
```

```{r, incluede=FALSE}
library(tidyverse)
library(kableExtra)
library(knitr)
table_format <- if (is_html_output()) {
  "html"
} else if (is_latex_output()) {
  "latex" 
}

fullw <- if (is_html_output()) TRUE else if (is_latex_output()) FALSE
fsize <- if (is_html_output()) 14 else if(is_latex_output()) 8


```

# Sample Design of the Study {#dis_muest}

The following section presents the description of the sample design of ELSOC, corresponding to the original sample and the refresh sample of the study. At the end of the section, the construction of the ELSOC weights is described.

## Sample Design Original Sample {#dis_muest_original}

The design of the original ELSOC sample aimed to reconcile the multiple research interests of the researchers associated with the Center. Among the most relevant considerations were the following:

1. A sample design that would allow combining the variables measured in the questionnaire with the spatial variables, recorded at the block and commune level, contained in the databases developed by the Territorial Intelligence Center (CIT) of the Universidad Aldolfo Ibáñez. Since CIT data is not available for all blocks in the country, particularly those located in rural areas, it was decided to include only urban areas in the sample. This consideration also coincided with the preferences of many researchers at the Center, who expressed being mainly interested in an urban sample.

2. Some researchers requested a design that would allow estimating multi-level (or hierarchical) models grouped by city and commune, and therefore, it was established that the sample should contain a sufficient number of cities and communes, as well as a sufficient number of cases within each city and commune, to allow such analysis (Snijders & Bosker, Chapter 10).

3. Other researchers were interested in comparing the inhabitants of the three largest cities in the country, which resulted in a non-proportional design that increased the number of respondents in the areas of Greater Valparaíso (cities of Viña del Mar and Valparaíso) and Greater Concepción (Concepción, Talcahuano, and others).

4. Finally, some researchers requested a design that would allow comparing respondents living in large and small cities, which favored increasing the sample size of households in small cities (Kish, 1965, Section 3.5), particularly those with between 30,000 and 100,000 inhabitants.

COES researchers worked with the person in charge of the sample design, Stephanie Eckman, to develop a design that could reasonably meet these needs and substantive interests. The final sample design of wave 1 of ELSOC COES provides adequate coverage of the largest cities in the country (Greater Santiago, Greater Valparaíso, and Greater Concepción), as well as smaller cities, and also ensures the representation of people in the north and south of the country. Overall, the sample design achieves an approximate representativeness of 77% of the total population of the country and 93% of the urban population. The following subsections detail the different steps of the sample design.

### Preparation of the Sampling Frame {-}

The sampling process for the original sample was based on the 2011 pre-census data, which was formatted by the CIT. Although the population counts from the 2012 census are not accurate, the pre-census work collecting information on households in all blocks is of good quality. The dataset contained a total of 155,757 blocks, but four different types were removed before selection began.

1. Following the analytical interests of the Center's researchers, only urban blocks were used. To determine which blocks were urban, the locality type coding (urban or rural) contained in the 2011 pre-census database was used. Consequently, 22,188 (14.2%) blocks were excluded at this step.

2. Again, based on the analytical interests of the Center's researchers, only blocks that had been previously geo-referenced by the CIT were retained for sampling. This implies that a total of 1,971 (1.5% of urban blocks) that were not geo-referenced were removed at this step.

3. Only blocks containing five or more households (according to the 2011 pre-census) were retained. 503 blocks (less than 1% of the remaining blocks after steps 1 and 2) did not meet this threshold and were eliminated.

4. Only blocks in cities with more than 10,000 people were eligible for selection. 10,238 blocks (7.8% of the remaining blocks) were excluded from the sampling frame.

Thus, the final sampling frame contains 120,857 blocks. The COES sample will represent only these blocks and not those that were excluded. Estimates derived from the sample data will apply only to this target population and should not be applied to the entire Chilean population. The selection process of respondents was developed in four stages, although a fifth stage was added during fieldwork.

### Stage 1: Selection of Cities {-#etapa1m1}

The universe of blocks (the 120,857 mentioned) was aggregated at the city level, resulting in 122 cities. The three largest cities (Greater Santiago, Viña del Mar - Valparaiso, and Concepción - Talcahuano) were selected with certainty. The remaining cities were stratified by population. Table \@ref(tab:tabla-estratos) shows the definitions of the strata and the population and sample sizes in each.

```{r tabla-estratos, echo=FALSE}

estratos <- data.frame(estrato = c('Great Santiago', 'Great Valparaíso', 'Great Concepción', 
                               'Large cities', 'Medium sized cities', 'Small cities'),
                   def_N = c(' ', ' ', ' ', '> 100 mil', '> 30 mil', '> 10 mil'),
                   N_ciudades = c(1, 1, 1, 18, 28, 73),
                   n_ciudades = c(1, 1, 1, 8, 10, 19),
                   N_norte = c(NA, NA, NA, 8, 15, 24),
                   n_norte = c(NA, NA, NA, 4, 6, 6),
                   N_sur = c(NA, NA, NA, 10, 13, 49),
                   n_sur = c(NA, NA, NA, 3, 3, 13)
                   )

estratos %>% 
  kbl(caption = 'Population by city and sample sizes, by stratum',
      align = c('l', rep('c', 7)),
      format = table_format,
      booktabs = TRUE,
      col.name = c('Stratum', 'Definition (N° inhabitants)', 'City population size', 'City sample size',
                   rep(c('Stratum population size', 'Stratum sample size'), 2))) %>% 
  kable_styling(full_width = FALSE,
                latex_options = c('HOLD_position', 'scale_down')) %>% 
  column_spec(2:8, width = "4em") %>% 
  kable_classic_2() %>% 
  add_header_above(c(" " = 1, " " = 3, "North Stratum" = 2, "South Stratum" = 2))



```

The strata of large cities, medium cities, and small cities were geographically stratified by North or South zone, to ensure that the sample contained cities from both the north and south of Chile. This results in a total of nine strata. The sample was allocated between the two areas in proportion to the size of their population in the universe. See Table \@ref(tab:tabla-estratos) for details on population and sample sizes in each of the north and south strata.

The selection of cities within each of these final strata was done proportionally to the size of the population of each city. This method gives higher probabilities of selection to larger cities.

The probability of selecting a city $i$ within stratum $h$ was:

$$\pi_i=\frac{(nc_h)(pop_i)}{\sum_h pop}$$

where $nc_h$ is the number of cities selected in stratum $h$ and $pop_i$ is the population of city $i$.

### Stage 2: Selection of Blocks (Manzanas) {-#etapa2m1}

The 40 selected cities contained 87,839 blocks. In the second stage, blocks were selected in each city with population proportional to size, where size was determined from the pre-census housing unit count. The selection was systematic: the list of blocks in the selected cities was ordered by sub-census district and block number to ensure that the selected blocks were spread throughout the city^[The block and census district numbers were provided by Matías Garretón, CIT researcher. Sub-census districts are smaller geographic units than the commune but larger than the blocks.].

Table \@ref(tab:tabla-ciudades) shows the number of blocks selected in each city, by stratum. The block sample was allocated disproportionately so that areas outside Santiago were over-represented relative to their size in the target population. Several COES researchers requested this allocation to ensure that the sample was diverse with respect to city size.

The probability of selecting a block $j$ in city $i$, conditional on the selection of the city, was:

$$\pi_{j|i}=\frac{(nb_i)(hu_j)}{\sum_i hu}$$

where $nb_i$ is the number of blocks selected in city $i$ and $hu_j$ is the population of city $i$.

```{r tabla-ciudades, echo=FALSE}

ciudades <- data.frame(estrato = c('Great Santiago', 'Great Valparaíso', 'Great Concepción', 
                               'Large cities', 'Medium sized cities', 'Small cities',
                               'Total'),
                   def_N = c(' ', ' ', ' ', '> 100 mil', '> 30 mil', '> 10 mil', ' '),
                   N_ciudades = c(1, 1, 1, 8, 10, 19, 40),
                   n_ciudades = c(200, 100, 100, 26, 25, 11, 27)) %>% 
  mutate(N_bloques = N_ciudades * n_ciudades)

ciudades %>% 
  kbl(caption = 'Distribution of cities and blocks by stratum',
      align = c('l', rep('c', 4)),
      format = table_format,
      booktabs = TRUE,
      col.name = c('Stratum', 'Definition (N° inhabitants)', 'City sample size', 'Number of blocks per city', 'Number of blocks')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = c('HOLD_position')) %>% 
  column_spec(2:4, width = "6em") %>% 
  kable_classic_2()
  

```

In 4 cities, some blocks were so large that they were selected with certainty. That is, the housing unit counts were greater than the selection interval and would be selected in any sample, and could even be selected twice. To avoid duplicate selections, these blocks were chosen first with certainty and then additional blocks were selected from the remaining ones for those cities, to reach the total desired sample size for the city (see Table \@ref(tab:tabla-ciudades). $\pi_{j|i}$ for these cities is 1.

The 1,067 blocks selected in the 40 chosen cities were listed in the field to carry out the selection of households with the most up-to-date information possible. The CIT provided maps of each selected block. CMD field staff visited each block in person and created a listing of all housing units in those blocks. The lists were carefully reviewed for any errors or duplicates.

During the listing process, the Microdata Center found that some blocks had more than 100 households, which makes the listing process excessively difficult. Consequently, these blocks were divided into sub-blocks of approximately equal size (40 to 50 households) and one was selected to be listed. Since the sub-blocks were created to be of similar sizes, they were selected based on equal probability. In total, 301 blocks were sub-sampled. The remaining blocks were not affected by this stage.

### Stage 3: Selection of Households {-#etapa3m1}

The number of households selected in each block varied by stratum, as shown in Table \@ref(tab:bloques). This design resulted in 4,001 housing units, aiming to obtain approximately 3,000 complete interviews, assuming a response rate of 75% for all strata.

```{r bloques, echo=FALSE, message=FALSE, warning=FALSE}

bloques <- data.frame(estrato = c('Great Santiago', 'Great Valparaíso', 'Great Concepción', 
                               'Large cities', 'Medium sized cities', 'Small cities',
                               'Total'),
                   def_N = c(' ', ' ', ' ', '> 100 mil', '> 30 mil', '> 10 mil', ' '),
                   n_ciudades = c(5, 5, 5, 3, 3, 3, 4001))

bloques %>% 
  kbl(caption = 'Distribution of households by blocks, by stratum',
      align = c('l', rep('c', 2)),
      format = table_format,
      booktabs = TRUE,
      col.name = c('Stratum', 'Definition (N° inhabitants)', 'Number of households per block')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = c('HOLD_position')) %>% 
  kable_classic_2()

```

A simple random sample of households was conducted in each block. The combination of population proportional to size sampling in the first two stages and simple random sampling in the third and fourth stages resulted in a sample of households with approximately equal probability within each of the nine strata.

The probability of selecting a household $k$ in block $j$ in city $i$ and stratum $h$ was:

$$\pi_{k|j,i}=\frac{nh_j}{NH_j}$$

where $nh_j$ is the number of households selected in block $j$, and $NH_j$ corresponds to the number of households listed in block $j$.

### Stage 4: Selection of Individuals {-#etapa4m1}

Interviewers visited each selected household and attempted to conduct the interview. The first step in the interview process was to identify the target respondent. When there was more than one adult in the household, one was selected using a simple random sample, using a Kish table.

The probability of selecting an individual in household $k$ was:

$$\pi_{l|k,j,i}=\frac{1}{NP_j}$$

where $NP_j$ is the number of adults (over 18 years old and under 75 years old) living in household $j$.

### Stage 5: Increasing the Sample Size {-#etapa5m1}

During the fieldwork of the first wave (2016), it was observed that the assumption of a 75% response rate for all strata was incorrect. First, the overall response rate was lower than 75%, and second, there was significant heterogeneity in response rates between regions. Due to this, it was decided to increase the number of households per block to effectively achieve the 3,000 interviews.

This increase in the number of households per block has a limited effect on the probability of selection of each household. It only affects the probability calculated in Stage 3, as the number of available households is smaller, but there are no changes in the probabilities calculated in Stages 1 and 2. This occurs because the selected blocks (in Stage 2) were used, and no new blocks were introduced.

During this process, a total of 1,082 new households were added to the study sample, located within the selected blocks. The allocation of these new households was not uniform across all blocks in the country. Instead, they were concentrated in four regions: Coquimbo, O'Higgins, Metropolitan, and Biobío regions, where interviewers had more difficulty contacting respondents. Table \@ref(tab:tabla-aumento) details the communes where the number of households was increased compared to the initial design, along with the total number of households added per block.

```{r tabla-aumento, echo=FALSE, message=FALSE, warning=FALSE}

tabla_aumento_muestra <- data.frame(region = rep('', 3),
          comuna = c('Coquimbo', 'La Serena', 'Salamanca'),
          viv_agregadas = c(24,28,22),
          viv_agregadas_bloque = c(2,2,2)) %>% 
  add_row(region = rep('', 3),
          comuna = c('Doñihue', 'Rancagua', 'Santa Cruz'),
          viv_agregadas = c(10,42,11),
          viv_agregadas_bloque = c(1,2,1)) %>% 
  add_row(region = rep('', 6),
          comuna = c('Chiguayante', 'Concepción', 'Coronel', 'Penco', 'Quillón', 'San Pedro de la Paz'),
          viv_agregadas = c(24,75,11,4,6,28),
          viv_agregadas_bloque = c(3,3,1,1,1,2)) %>% 
  add_row(region = rep('', 34),
          comuna = c('Cerrillos', 'Colina', 'Curacaví', 'El Bosque', 'Estación Central', 'Huechuraba', 'Independencia', 'Isla de Maipo ', 'La Cisterna', 'La Florida', 'La Granja', 'La Pintana', 'La Reina', 'Las Condes', 'Lo Barnechea', 'Lo Espejo', 'Lo Prado',  'Macul', 'Maipú', 'Ñuñoa', 'Padre Hurtado', 'Pedro Aguirre Cerda' , 'Peñaflor' ,'Peñalolén', 'Providencia', 'Pudahuel', 'Puente Alto','Quilicura', 'San Bernardo', 'San Joaquín', 'San Miguel', 'San Ramón', 'Santiago', 'Vitacura'),
          viv_agregadas = c(9,12,14,8,12,6,6,39,9,24,6,12,9,33,9,6,6,8,32,16,6,6,30,14,7,14,32,12,16,6,9,6,120,9),
                  viv_agregadas_bloque = c(3,3,2,2,3,2,2,3,3,2,2,2,3,3,3,2,2,2,2,2,3,2,2,2,3,2,2,2,2,2,3,2,3,3))


tabla_aumento_muestra %>% 
  kbl(caption = 'Number of households added to the sample by region and commune',
      align = c('l', 'l', 'c', 'c'),
      format = table_format,
      booktabs = TRUE,
      longtable = TRUE,
      col.name = c('Region', 'Commune', 'Total households added', 'Households added per block')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = c('repeat_header')) %>% 
  kable_classic_2() %>% 
  pack_rows('Coquimbo', 1, 3) %>% 
  pack_rows('O’Higgins', 4, 6) %>% 
  pack_rows('Biobío', 7, 12) %>% 
  pack_rows('Metropolitana', 13, 46)


```

### Sample Design for Successive Waves in Original Sample {-}

The sample design for successive waves is equivalent to the original sample design. The households added and the respondents selected during 2016, as well as those added during [Stage 5](#etapa5), were re-interviewed in the follow-up waves.

## Sample Design Refresh Sample {#dis_muest_refresco}

Starting in 2018, it was decided to include a refresh sample to counteract the effects of attrition (see [Survey Attrition](#atricion)). The design is explained below.

### Preparation of the Sampling Frame for the Refresh Sample {-}

Since the 2017 Census databases were not yet available to be used as a housing sampling frame at the time of preparing the sampling frame for the definition of the Refresh sample, it was decided to use the household listing conducted by CMD during 2016 as the sampling frame.

The listing corresponds to a list of households surveyed in person in the first round of the ELSOC survey (Original sample), so it consists of the blocks (blocks) selected for the Original sample survey. This includes around 40,000 households in 1,067 blocks (see [Stage 2: Selection of Blocks](#etapa2)).

Since the listing is two years old at the time of generating the sample design for the refresh sample, coverage problems may arise in the listed blocks. Additionally, there are 179 segments with fewer than 20 listed households; these segments were not considered for the selection of the new sample. If a selected segment was not residential, a replacement would be considered in the sample selection.

Again, the selection process of households and respondents was developed in 4 stages:

### Stage 1: Selection of Cities {-#etapa1m2}

For the selection of Cities, it was decided to use the same cities as the Original sample, present in the listing conducted by CMD. Therefore, the sample design of the Refresh Sample is equivalent to the sample design of the Original Sample up to Stage 1 of city selection.

### Stage 2: Selection of Blocks {-#etapa2m2}

The blocks were selected following a simple random sample among those present in the CMD listing and selected in the Original Sample. This decision was made considering that it simplifies the information collection process, as it does not alter the blocks to be visited (thus increasing the productivity of the collection), and allows maintaining similar representativeness concerning the Original sample.

However, before the selection, blocks with fewer than 20 households or those that were part of the pilot survey for the 2018 refresh sample are discarded. This means that the universe of available blocks is reduced from 1047 to 822. Finally, since the Refresh Sample has a target size of 1,400 individuals, 402 blocks are selected.

Table \@ref(tab:seleccion-bloques-m2) presents the listed blocks, available blocks, and selected blocks, by region:

```{r seleccion-bloques-m2, echo=FALSE}

bloques_refresco <- data.frame(region = c('Tarapacá', 'Antofagasta', 'Atacama', 'Coquimbo', 'Valparaíso', 'Metropolitana', 'O\'Higgins', 'Del Maule', 'Biobío', 'La Araucanía', 'Los Ríos', 'Los Lagos', 'Aysén', 'Total'),
                               segmentos_empadronamiento = c(23, 24, 51, 73, 159, 283, 47, 92, 143, 94, 22, 25, 11, 1047),
                               segmentos_no_considerados = c(3, 3, 18, 22, 28, 32,7, 24, 34, 24, 6, 3, 1, 205),
                               segmentos_piloto = c(0, 0, 0, 0, 0, 20, 0, 0, 0, 0, 0, 0, 0, 20),
                               segmentos_disponibles = c(20, 21, 33, 51, 131, 231, 40, 68, 109, 70, 16, 22, 10, 822),
                               segmentos_seleccionados = c(8, 9, 18, 23, 70, 100, 16, 32, 64, 38, 9, 10, 5, 402))

bloques_refresco %>% 
  kbl(caption = 'Blocks available and selected, according to region',
      align = c('l', rep('c', 5)),
      format = table_format,
      booktabs = TRUE,
      col.name = c('Region', 'Registered blocks', 'Blocks < 20 dwellings', '2018 Pilot Blocks', 'Available blocks', 'Selected blocks')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = c('HOLD_position')) %>% 
  column_spec(2:6, width = "5em") %>% 
  kable_classic_2()


  
```

### Stage 3: Selection of Households {-#etapa3refresco}

The households are selected with a simple random sample among the households in the selected blocks. Since it is defined to obtain 1,400 surveys, it is necessary to select an average of 3.5 households per block to meet the target sample. For this, in those blocks that have more than 20 households, between 4 and 6 households per block are selected, depending on the size of the strata and blocks; and in those blocks with more than 50 households, another of the sub-blocks generated in the listing process is used, ensuring that this sub-block is not in the same building already selected.

The probability of selecting the same household in the block is different from zero, so a replacement will be generated for that unit using the same random selection method, for which the selected sample must be verified with the panel sample already defined in 2016.

### Stage 4: Selection of Individuals {-#etapa4refresco}

As in the Original Sample, interviewers visit each selected household and attempt to conduct the interview. The first step in the interview process is to identify the target respondent, who is selected using a simple random sample among the adults between 18 and 75 years old living in the household.

## Weight Design {#dis_ponderadores}

Considering that from the interviews actually conducted, it is desired to extrapolate to the respective target population, it is necessary to adequately weight each respondent according to their representation in the target population. For this, in each wave of the study, weights are provided that allow adjusting for differences in demographic attributes of the ELSOC sample concerning the target population. The weights included in the ELSOC database are the product of three factors: A design weight, a non-response adjustment, and a post-stratification adjustment.

The process of developing these weights is described below^[These weights can be understood as "cross-sectional weights," as they ignore the problems derived from their longitudinal nature.] to achieve a correspondence between the sample and the target population.

### Design Weights

The first step in constructing the ELSOC weights is to consider the inverse of the probability of selecting an individual into the sample. In this case, the probability of selecting individual $i$ from household $j$ belonging to block $l$ within stratum $k$, $P_{ijlk}$ is given by:

$$P_{ijlk} = \pi_{i|jlk}\pi_{j|lk}\pi_{lk}$$

Where:

* $\pi_{i|jlk}$ is the probability that individual $i$ is selected in the sample given that the household where they live and the block where it is located were selected, within stratum $k$.
* $\pi_{_j|lk}$ is the probability that household $j$ is selected in the sample given that block $l$ (which contains household $j$) was selected.
* $\pi_{lk}$ is the probability that block $l$ of stratum $k$ is selected in the sample.

The design weight $w_{ijlk}$ is defined as the inverse of the probability of selection of each individual in the sample:

$$w_{ijlk}=\frac{1}{ P_{ijlk}}$$

The probabilities that make up the selection probability are calculated by:

$$\pi_{lk} = n_k\frac{M_{lk}}{M_k}$$

$$\pi_{j|lk} = \frac{m_{lk}}{M'_{lk}}$$

$$\pi_{i|jlk} = \frac{1}{N_{jlk}}$$

Where $n_k$ is the number of blocks to be selected from stratum $k$, $M_{lk}$ is the number of households in block $l$ of stratum $k$, $M_k$ is the total number of households in stratum $k$. $m_{lk}$ is the number of households to be surveyed within block $l$, $M'_k$ is the updated number of households in block $l$ post-listing, and $N_{jlk}$ is the number of people in the target population living in household $j$ of block $l$ of stratum $k$.

### Non-Response Adjustments

The previous weight is then adjusted based on the initial non-response of the block in which each individual is located. In this way, more weight is given to observations in blocks where few people respond.

This adjustment is made based on a non-response factor for block $f_l$, and is estimated as the inverse of the proportion of households within block $l$ that respond to the survey.

$$f_l=\frac{m_l}{resp_l}$$

Where $m_l$ is the number of households selected to be in the sample, and $resp_l$ is the number of selected households that participate in ELSOC.

### Post-Stratification

Post-stratification is an adjustment commonly made in surveys to incorporate known information about the population into the analysis of survey data. Its main objective is that the distribution of weighted estimators can reproduce the population distribution in key variables such as regional totals and by sex.

The non-response adjustment can be thought of as a factor that adjusts:

$$ps=N_j/\hat{N_j}$$
Where $N_j$ is the total number of people in the population within stratum j (e.g., total number of women in the population), and $\hat N_j$ is the estimated total of that total from the sample data.

Finally, the cross-sectional weights of ELSOC have the form:

$$w=w^{dis}*f_l*ps$$
That is, they are a combination of the probability of inclusion of individuals, a non-response adjustment in the block where they reside, and a post-stratification adjustment.

### Longitudinal Weights

The use of weights in statistical analyses has multiple motivations, one of the most important being the interest in avoiding biases due to the sample design of the study. Given the longitudinal nature of ELSOC, it is particularly important to control for attrition biases that arise from the existence of systematic patterns in non-participation in the study after the first wave. Another relevant motivation for the use of weighted estimators is their ability to incorporate known information about the target population in the construction of the weights. To achieve this, the weights are adjusted so that they can reproduce population totals or proportions.

A complete report on the generation and use of longitudinal weights for ELSOC data has been generated, which can be reviewed on the site [https://pondedradores-long-elsoc.netlify.app](https://pondedradores-long-elsoc.netlify.app).

\newpage
