```{r}
library(tidyverse)
library(kableExtra)
```

# Instrument Design {#dis_ins}

The information collection instrument consists of a structured questionnaire (survey type) applied to all participants. The questionnaires for each measurement can be found at [**coes.cl/cuestionarios**](https://coes.cl/cuestionarios/).

The questionnaire was designed to measure a series of conceptually relevant aspects that allow characterizing the levels of conflict and social cohesion in Chile, emphasizing its evolution over time. The main topics of analytical interest addressed by the survey correspond to the modules in which it is structured:

- Citizenship and Democracy
- Social Networks and Inter-group Interactions
- Legitimacy and Social Inequality
- Social Conflict
- Neighborhood Dimension
- Health and Well-being
- Socio-demographic Characterization
- Gender (added to the survey since 2019)

The application of the questionnaire was originally carried out using the **CAPI** system (Computer-Assisted Personal Interviews). During the 2021 wave, it was changed to the **CATI** system (Computer-Assisted Telephone Interviews) due to the COVID-19 contingency (more details in the section [Questionnaire 2021: data collection during the COVID-19 pandemic](#instrumento-covid)). During the 2022 wave, it switched to a Mixed format (or Mix-mode), combining face-to-face (CAPI) and telephone (CATI) formats (more details in the section [Change to mixed-mode data collection](#instrumento-mixmode)).

## Questionnaire Design {#dis-cuestionario}

The original ELSOC questionnaire design was developed during 2015 and covered the planning of the 2016 and 2017 waves. Differences between subsequent questionnaires are described in the section [Changes in the Questionnaire between waves](#dif-cuest-olas).

Most of the questions, scales, and/or items included in the ELSOC modules come from other public opinion studies, research in psychology, sociology, economics, political science conducted by national and international academics, and social surveys conducted in Chile and other countries. Additionally, some scales were developed by COES team members and/or adapted from their previous studies. The questionnaire was designed by leveraging the main technical recommendations and the state of the art in the various areas of study included.

To meet the quality criteria and standards for questionnaires and to reconcile the multiplicity of research agendas developed by COES, a working procedure was established for its construction. This process was carried out during 2015 in different phases:

1. COES-affiliated researchers were asked to propose research projects that include a theoretical framework and hypotheses supporting the proposed scales to be included in the questionnaire. Proposals could be submitted individually or collectively, with no restrictions on the number of items to propose. The only requirement was an explicit theoretical foundation involving longitudinal hypotheses and an operationalization of the constructs to be measured in the proposed items.

2. The previous procedure involved receiving and organizing a large number of research agendas, resulting in more than 750 items (unique questions or part of a scale). The ELSOC team systematized the received proposals, classifying the items into thematic areas. Mechanisms were designed to reduce the number of items, and meetings were coordinated among COES researchers to decide on them. The selection criteria were mainly theoretical, prioritizing fundamental questions for the longitudinal analysis of research projects and their alignment with COES's substantive agenda. Practical criteria were also included, such as eliminating identical or very similar items, previous evidence on item quality (previous surveys, especially the cross-sectional survey conducted by COES in 2014), and the overall study design (sampling design, unit of analysis, type of informant, etc.).

3. After selecting the most relevant scales for each proposed topic, a pilot study of the developed questionnaire was conducted. This involved piloting 430 items, with the MicroData Center of the University of Chile responsible for its execution. Based on the pilot results, the panel team made adjustments to some items and developed a proposal to reduce items, which was evaluated with the Center's researchers.

4. The final stage of questionnaire adjustment focused on identifying the most conceptually relevant aspects for COES and weighing the methodological requirements for their empirical evaluation. Items were classified according to the number of required measurements, distinguishing between permanent items (measured in all waves, as they constitute the study's analytical core) and intercalated items (measured only once or every other wave).

The final version of the information collection instrument consists of a structured questionnaire (survey type) applied face-to-face to all participants annually, with a total of 326 items or questions.

### Changes in the Questionnaire between waves {-#dif-cuest-olas}

As described in the previous section, ELSOC questions are classified into permanent and intercalated items, as some are not measured in all waves. Therefore, the implemented questionnaires differ between the various waves of the survey.

Additionally, the implemented questionnaires have undergone modifications that were not originally planned due to a constant evaluation process of the applied instrument's quality, which may result in modifying or eliminating questions; the existence of contingencies that, given the study's nature, motivate the inclusion of new items; or the emergence of new research agendas relevant to ELSOC.

Another factor that impacted the implemented questionnaire is the average interview duration. During 2016, the average duration per respondent was 55 minutes. CMD recommended the ELSOC team reduce the interview time to an average of 45 minutes, as longer durations lead to greater fatigue for interviewers and respondents, negatively affecting data collection quality. Therefore, it was decided to shorten the study's length. The ELSOC team, together with COES researchers, defined various adjustments to reduce the questionnaire's length. While the 2016 wave questionnaire contained 326 items, the 2017 instrument included 309. Efforts have been made to shorten the questionnaire in subsequent waves.

Table \@ref(tab:tabla-items-olas) shows the number of items measured in each wave of the study in the Original and Refresh samples:

```{r tabla-items-olas}

items_olas <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'), 
                         sheet = 'n_ola_muestra') %>% 
  janitor::clean_names() %>% 
  select(ola, muestra_original, muestra_refresco)

items_olas %>% 
  kbl(caption = 'Number of items per sample, by wave',
      align = c('l', 'c', 'c'),
      booktabs = TRUE,
      col.name = c('Wave', 'Items in original sample', 'Items in refresh sample')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2()

```

Table \@ref(tab:tabla-items-olas-modulos) shows the distribution by questionnaire module:

```{r tabla-items-olas-modulos}

items_olas_modulos <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'), 
                         sheet = 'n_ola_muestra_modulo') %>% 
  janitor::clean_names()

items_olas_modulos %>% 
  dplyr::select(-muestra) %>% 
  kbl(caption = 'Number of items per module, by sample and wave',
      align = c('l', 'c', 'c', 'c', 'c', 'c', 'c'),
      booktabs = TRUE,
      col.name = c('Module', '2016', '2017', '2018', '2019', '2021', '2022', '2023')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2() %>% 
  pack_rows('Original Sample', 1, 8) %>%
  pack_rows('Refresh Sample', 9, 16)

```

### Differences in Questionnaires between Samples {-#dif-cuest-muestras}

Starting in 2018, the Refresh Sample was implemented to complement and compensate for the accumulated attrition of the Original Sample. The refresh sample questionnaire is largely similar to the original sample. However, given the wide range of research agendas and topics relevant to ELSOC, the inclusion of questions present only in one of the two samples has been allowed to increase the number of covered topics while maintaining a more limited number of questions asked to each respondent.

Table \@ref(tab:tabla-items-compartidos) shows the number of items shared between both samples by wave, as well as those items present exclusively in the original or refresh sample.

```{r tabla-items-compartidos}

items_compartidos <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'), 
                         sheet = 'n_ola_muestra_traslape') %>% 
  janitor::clean_names() 

items_compartidos %>% 
  kbl(caption = 'Number of items per sample, by presence in Original and Refresh samples',
      align = c('l', 'c', 'c', 'c', 'c'),
      booktabs = TRUE,
      col.name = c('Wave', 'Only in original sample', 'Only in refresh sample', 
                   'Both samples', 'Total')) %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2() %>% 
  column_spec(2:5, width = "7em")

  
```
### Questionnaire 2021: data collection during the COVID-19 pandemic {-#instrumento-covid}

The health crisis that began in 2020 due to COVID-19 has posed significant challenges to statistical systems in general and public opinion surveys in particular. The International Labour Organization (ILO, 2020) reports that most national statistical offices have experienced a significant impact on their operations, particularly those conducted in person. To address these issues, several organizations have had to transform their face-to-face operations into telephone or web surveys.

In this context, to ensure the feasibility of the ELSOC 2021 survey data production process and the safety of both respondents and interviewers, the ELSOC executive team, together with professionals from the Microdata Center of the University of Chile, decided to adopt a series of measures that temporarily transitioned to a telephone questionnaire format (CATI mode).

Since an increase in application times was expected when switching to the telephone format, the questionnaire application was divided into two calls, called Calls A and B, respectively, with the aim of each having a maximum duration of 30 minutes to reduce the interview application time and avoid respondent and interviewer fatigue.

To evaluate the challenges and methodological change of the application, a pilot survey was conducted between December 2 and 21, 2020, which was relevant for evaluating both the technical and methodological aspects associated with the application change. In this process, it was found that the average application time exceeded 30 minutes per call, so it was decided to:

1. Reduce the number of questionnaire items
2. Adjust questions that generated complications based on the pilot test findings

The following criteria were defined to reduce the questionnaire size: 1) reduce dimensions that have been significantly constrained by quarantines, such as political participation and social interaction; 2) prioritize items based on technical consistency and/or alignment with COES objectives; 3) prioritize items with fewer than three measurements throughout the study; and 4) maintain critical socioeconomic and health items to adequately assess the pandemic and quarantine impacts.

Regarding the allocation of questions between Calls A and B, the main criteria used were: 1) maintain the quality of the survey flow by keeping questions on similar topics in the same call and ensuring that variable batteries sharing the same initial header remained together; and 2) ask essential items related to the pandemic, quarantines, and social unrest immediately to monitor and characterize respondents during this period.

Table \@ref(tab:tabla-llamados) presents the distribution of items by module for Calls A and B, distributed by sample type.

```{r tabla-llamados}

items_llamado_2021 <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'), 
                         sheet = 'n_llamados_2021') %>% 
  janitor::clean_names()

items_llamado_2021 %>% 
  kbl(caption = 'Distribution of items by call and sample, Wave 2021',
      align = c('l', 'c', 'c', 'c', 'c'),
      booktabs = TRUE,
      col.name = c('Wave', rep(c('Call A', 'Call B'), 2))) %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2() %>% 
  add_header_above(c(" "= 1, "Original Sample" = 2, "Refresh Sample" = 2))

  
```

Secondly, and linked to the above, variables with more than five response categories that generated greater complications during the pilot tests of this survey were adjusted. Below are the items that had their response alternatives reduced and the criteria adopted for each case:
 
 - Distant Networks Battery [r01, r02, and r04]: This battery asks about the number of people in different occupations and social groups that the respondent knows. Originally, this battery of questions has 7 values (1. None; 2. One; 3. Between 2 and 4; 4. Between 5 and 7; 5. Between 8 and 10; 6. Between 11 and 15; 7. 16 or more). Since the response ranges do not follow a clear pattern, it was decided to ask respondents for the exact number of acquaintances.
 
  - Respondent's weight [s07]: This variable has 9 response ranges, which were established in 5 ranges, based on the weight quintiles reported by ELSOC in the 2018 wave in variable s06.
  
 - Income variables in ranges [m14 and m30]: Variable m14 has 16 response ranges, while variable m30 has 30 response ranges. Both variables were established in 5 ranges, based on the income quintiles presented by the Casen 2017 survey.

### Questionnaire 2022: change to mixed-mode data collection {-#instrumento-mixmode}

Starting from the sixth wave of ELSOC, the **mix-mode** method or **mixed** methodology was used, combining CAPI and CATI formats to conduct the surveys, thus optimizing the data collection process.

The main data collection methodology, however, continues to be the CAPI modality: the survey was conducted in CATI mode only when the informant declared that they could not or did not want to answer the survey in person. As a result, during 2022, 13.3% of the original sample and 15.8% of the refresh sample were conducted by telephone format (see \@ref(tab:tabla-freq-mixmode)).

```{r tabla-freq-mixmode}

freq_mixmode <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'),
                                sheet = 'n_mixmode') %>% 
  mutate(porc_cati_m1 = scales::percent(m1_cati / (m1_cati + m1_capi), .1),
         porc_cati_m2 = scales::percent(m2_cati / (m2_cati + m2_capi), .1)) %>% 
  select(ola, m1_capi, m1_cati, porc_cati_m1, m2_capi, m2_cati, porc_cati_m2)

freq_mixmode %>% 
  kbl(align = rep('c', 7),
      booktabs = TRUE,
      col.name = c('Measurement', rep(c('CAPI', 'CATI', '% CATI'), 2)),
      caption = 'Number of surveys by interview format, by wave and sample') %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2() %>% 
  add_header_above(c(" " = 1, "Original Sample" = 3, "Refresh Sample" = 3))

```

As in the fifth wave data collection, the telephone modality considered conducting the interview in 2 calls to avoid exhausting the informant and thus prevent loss of data quality.

Comparative evidence reveals that the mixed-mode interview modality increases coverage; and reduces measurement errors (Dillman et al., 2014^[Dillman, D. A., Smyth, J. D., & Christian, L. M. (2014). Internet, phone, mail, and mixed-mode surveys: The tailored design method. John Wiley & Sons.]), and representation biases of lower-income people and those with greater material deprivations (Lipps, 2016^[Lipps, O. (2016). Non-observation bias in an address-register-based CATI/CAPI mixed mode survey. methods, data, analyses, 10(1), 20.]), associated with the data collection process. Other studies show that there are no significant differences in the degree of confidence and stability of estimates when comparing CAPI and Mixed collection methods (Cernat, 2015^[Cernat, A. (2015). The impact of mixing modes on reliability in longitudinal studies. Sociological Methods & Research, 44(3), 427-457.]), nor in response rates (Bianchi, Biffignandi, and Lynn (2016)^[Bianchi, A., Biffignandi, S., & Lynn, P. (2016). Web-CAPI sequential mixed-mode design in a longitudinal survey: effects on participation rates, sample composition and costs (No. 2016-08). Understanding Society at the Institute for Social and Economic Research.]).

Given the international evidence, it can be considered that there are no differences in the degree of confidence and stability associated with the different collection methods, and no differences in the characteristics of those who decide to participate should be expected. On the contrary, this new design could reduce representation bias, decrease attrition, and increase ELSOC's geographic coverage.

### Questionnaire 2023 {-#instrumento-normal}

```{r tabla-freq-normal}

items_llamado_2023 <- readxl::read_xlsx(path = file.path('inputs', 'datos_manual.xlsx'), 
                         sheet = 'n_llamados_2023') %>% 
  janitor::clean_names()

items_llamado_2023 %>% 
  kbl(caption = 'Distribution of items by call and sample, Wave 2023',
      align = c('l', 'c', 'c', 'c', 'c'),
      booktabs = TRUE,
      col.name = c('Wave', rep(c('Call A', 'Call B'), 2))) %>% 
  kable_styling(full_width = FALSE,
                latex_options = 'HOLD_position') %>% 
  kable_classic_2() %>% 
  add_header_above(c(" "= 1, "Original Sample" = 2, "Refresh Sample" = 2))

```


## List of variables

Given the large number of items measured over time, to examine in detail the differences between the applied instruments, interested parties are suggested to review the **Questionnaires** for each measurement (available at [**coes.cl/cuestionarios**](https://coes.cl/cuestionarios/)). Similarly, it is suggested to review the [**List of variables**](https://coes.cl/wp-content/uploads/Listado-de-Variables-ELSOC-2.xlsx) and the [**Codebook**](#libro-codigos), which include all items measured in ELSOC, with detailed information on the waves in which each question was measured, as well as modifications and other relevant information for the use of ELSOC data.

\newpage

