# Ethics

## Ethics Committee Approval

The Longitudinal Study on Conflict and Social Cohesion in Chile (ELSOC) has ethical approval for all its waves from the [Scientific Ethics Committee of Social Sciences, Arts, and Humanities of the Pontifical Catholic University of Chile](http://eticayseguridad.uc.cl/comite-etico-cientifico-en-ciencias-sociales-artes-y-humanidades.html) dated June 8, 2016, with **Protocol ID 160129004**.

The Ethics Committee approval was renewed on January 6, 2021.

## Informed Consent

The study participants signed a Written Informed Consent detailing the study's objectives, their participation, and how the information and data provided will be handled and safeguarded.

The participants of the Original Sample signed the Informed Consent in 2016, while the participants of the Refresh Sample signed it in 2018. Additionally, due to the change in modality during the 2021 wave data collection (see [2021 Questionnaire: data collection during the COVID-19 pandemic](#instrumento-covid)), participants from both samples were asked to sign an Appendix to the Informed Consent, detailing the changes made during that wave.

```{r, results = 'asis'}
  if (is_html_output()) {
    
    cat('The Informed Consents signed by participants in 2016 (Original Sample), 2018 (Refreshed Sample) and 2021 (Appendix) are available below:')
  }
```


```{r}

if (is_html_output()) {
  xfun::embed_file(file.path('inputs', 'consentimiento_informado_panel_coes_original.pdf'), text = 'Informed Consent 2016 (Original Sample)')
}

```


```{r}

if (is_html_output()) {
  xfun::embed_file(file.path('inputs', 'consentimiento_informado_panel_coes_refresco.pdf'), text = 'Informed Consent 2018 (Refrsco Sample)')
}

```


```{r}

if (is_html_output()) {
  xfun::embed_file(file.path('inputs', 'consentimiento_informado_panel_coes_apendice.pdf'), text = 'Informed Consent 2021 (Appendix)')
}

```

\newpage
